Demand Forecasting of Individual Probability Density Functions with Machine Learning

### 2020-12-23 LaTex/paper.tex
final review before resubmit

Demand forecasting is a central component of the replenishment process for retailers, as it provides crucial input for subsequent decision making like ordering processes. In contrast to point estimates, such as the conditional mean of the underlying probability distribution or confidence intervals, forecasting complete probability density functions allows to investigate the impact on operational metrics, which are important to define the business strategy, over the full range of the expected demand. Whereas metrics evaluating point estimates are widely used, methods for assessing the accuracy of predicted distributions are rare, and this work proposes new techniques for both qualitative and quantitative evaluation methods. Using the supervised machine learning method ``Cyclic Boosting'', complete individual probability density functions can be predicted such that each prediction is fully explainable. This is of particular importance for practitioners, as it allows to avoid ``black-box'' models and understand the contributing factors for each individual prediction. Another crucial aspect in terms of both explainability and generalizability of demand forecasting methods is the limitation of the influence of temporal confounding, which is prevalent in most state of the art approaches.

@@

% specific quantile
One of the main business operations of retailers is to ensure that sufficient goods are available in the store for customers to buy. This entails two parts: In a first step, we need to estimate the future customer demand to be able to judge how many goods we will need to order from the supplier or wholesaler. In the second step, aimed at inventory control, we need to place the actual orders, which is further complicated by the fact that multiple operational constraints, such as lot-sizes or delivery schedules, need to be included in the final ordering decision. Unfortunately, from the perspective of the retailer, the two parts, estimating the future demand and optimization of the order quantity, do not factorize into two separate issues, but are inextricably linked. This is because the business strategy depends on the complete chain of demand estimation and subsequent replenishment process to control the inventory. We can understand this in the following way: In the simplest case, a retailer might want to optimize their profit, or equivalently the cost, by reducing the impact of perishable items that have not been sold at the end of the selling period, as well as the lost revenue by not being able to meet the customer demand within this period. This is known as the ``newsvendor problem'' \cite{Edgeworth}, see, e.g., \cite{Khouja1999537} for a detailed review. Assuming linear overage ($h$) and underage ($b$) cost, the optimal quantile $q_{\mathrm{opt}} = {b}/{(b+h})$ of a known probability density function (PDF) describing demand as random variable can be calculated exactly. This quantile then corresponds to the point estimator corresponding to the numerical value of the demand prediction that can be used to determine the optimal order quantity.

% simulations
In reality, the situation is often more complex, as many factors may contribute to the cost function. Furthermore, the cost function may vary between individual stock-keeping units (SKU) and may also change with time. However, the general principle remains: If we have access to the relevant cost function, we can determine the optimal point estimator of the future demand from the associated PDF. Appendix \ref{sec:CostQuantile} demonstrates how the optimal point estimator can be calculated using the full PDF. For the case of a quadratic cost function, such as the mean squared error (MSE), or the linear cost function used in the newsvendor setting, the calculations are straight-forward. If the cost function becomes more complex, we may no longer be able to perform the calculations analytically, but will have to resort to numerical approaches. Furthermore, if we can predict more than one quantile, we can investigate the effect of choosing different points, and determine, for example, how sensitive operational metrics, such as stockout or waste rate of perishable goods, change when choosing a different quantile. This can be done using, for example, a business oriented simulation that starts from the quantiles of the predicted demand for each SKU to simulate the value of the business relevant KPIs for each setting, taking operational constraints into account. This can then be used, for example, to align the business strategy. Predicting the full demand PDF therefore allows the maximum flexibility for further optimizations, because we have direct access to all quantiles.

% PDF predictions
Due to its stochastic nature, demand is difficult to forecast: It depends on many influencing factors and, as already stated above, we can interpret it as a random variable with associated PDF. And as such, demand is neither identically and independently distributed (i.i.d.) between different products and sales locations, nor between different times. This means that, while we can generally assume that demand follows a specific type of PDF, its parameters are unique to the instance for which an estimate is required. For example, the PDF describing the demand of a particular product is specific to this product, date, and retail location for which the forecast is made. Furthermore, this distribution depends on a wide range of influencing factors, such as seasonal effects, sales price, local weather, etc. Therefore, in order to predict the expected demand for each SKU at each sales location and sales day, we need to forecast individual PDF predictions rather than globally adjusted parameters for groups of SKUs or locations or sales days.

% temporal confounding
If we account for the cases where demand exceeds the current stock level, resulting in out-of-stock situations, the actual demand in the past can be approximated by realized sales. We can then understand the time-ordered observations of sales events as a time series of a random variable. It is therefore tempting to use established time series regression models to predict the future values of the expected demand. However, if we wish to use the predicted values for anything but the prediction itself, we have to be careful to avoid, or at least limit as much as possible, temporal confounding, which emerges if some external variable $X$ causally influences several time steps $t$ of the observed variables $Y$ used in the time series, say $Y_{t}$ and $Y_{t-\tau}$ with lag $\tau$. In the case of a retailer, $Y_{t}$ is the demand for a specific product at a given sales location for a specific day, and $X$ can be, for example, the price of the product which we can set as part of our business strategy or change according to some promotion. This dependency creates a backdoor path between $Y_{t-\tau}$ and $Y_{t}$, that in turn leads to a spurious autocorrelation between the values of $Y$ at different times $t$, if we cannot control for the confounder $X$. This is particularly important for the case of demand forecasting, since customer demand is not a natural phenomenon we observe as an outsider, but retailers actively influence the customer demand  through a wide range of interventions, such as a given assortment choice, promotions, advertisements, coupons, etc. If we cannot disentangle the fundamental causal dependencies from the spurious autocorrelations, understanding the influences that lead to a specific prediction becomes very challenging, even if the underlying prediction method itself is fully explainable. Moreover, due to this disguising of actual causal dependencies, reliance on temporal confounding is also detrimental in terms of generalizability of the forecasting model. The issue of temporal confounding, together with a proposed solution, is discussed in more detail in sec. \ref{sec:ts_confounding}.

% ML
Demand forecasting for retailers often requires to predict millions of product-location combinations for each SKU each day. Univariate time series forecasting methods operate on each of these many individual time series separately. On the other hand, supervised machine learning (ML) algorithms optimize all individual time series together, which can enhance the generalizability of the method by exploiting common aspects between the various time series. This can lead to a drastic reduction in variance of the model, and in turn to an improvement of the forecast quality of each of the individual time series. Furthermore, machine learning provides a straight-forward way to include the many exogenous variables influencing demand as features (or covariates) of the model. This can also help to improve the forecast quality compared to methods mainly relying on autocorrelation of the historic sales.

Hence, in order for a retailer to be able to use demand forecasts most effectively, we need to be able to:
\begin{itemize}
	\item {Predict the future demands on SKU level at required temporal granularity as complete individual PDFs.}
	\item {Avoid temporal confounding in the model used for the prediction of the future demand.}
	\item {Evaluate that the functional form of the PDF is correct.}
	\item {Evaluate that the shape of the PDF is correct and that each part of the PDF models the observed data appropriately.}
	\item {Ideally, each prediction is fully explainable, enabling operational experts to exactly understand what contributed to it.}
\end{itemize}

The remainder of the paper is organized as follows: We first review the relevant literature and existing work in sec. \ref{sec:LitRev}. We then describe our method to predict individual negative binomial PDFs using  a parametric approach including two distinct machine learning models for mean of variance in sec. \ref{sec:pdfEstimation}. After that, we describe novel techniques for the qualitative and quantitative evaluation of PDF predictions in sec. \ref{sec:pdfEvaluation}. This is followed by a detailed discussion on temporal confounding in sec. \ref{sec:ts_confounding}. And finally, we present a demand forecasting example to show an application of our methods in sec. \ref{sec:example}.

@@

% univariate time series methods
A popular approach is to treat the demand as a time series, i.e., as a time-ordered sequence of observed sales events. In this setting, the future demand can be predicted as a point estimate using auto-regressive integrated moving average (ARIMA) \cite{BoxJenkins} or exponential smoothing \cite{croston1972forecasting, holt1957forecasting, Brown1963, Gardner1985} approaches. See also \cite{deGooijerHyndman2006} for a comprehensive overview over traditional time series forecasting methods. Essentially, these methods extrapolate the future values by building some form of regression model based on past observations, i.e., $y_{t+1} = f(y_t, y_{t-1}, \ldots)$, introducing dedicated functional components for trends, seasonality or external factors in various ways, depending on the exact modeling approach taken. The main underlying concept behind these approaches is to exploit the autocorrelation between values of the variable $y_t$ that is to be forecasted with some lag $\tau$ between subsequent events. Intuitively, we assume that  the sequence of events observed so far is a good predictor of future events. Examples for these methods in the context of demand forecasting can be found in, e.g., \cite{huber2017cluster, kalchschmidt2006forecasting, fattah2018forecasting, permatasari2018sales}.  A different approach is to use a concept from engineering, the Kalman filter \cite{kalman1960new}, and apply it to statistical forecasting \cite{morrison1977kalman}. Examples for this method used to estimate demand are, e.g., \cite{mitropoulos1980using, tegene1991kalman, kandananond2014applying,jacobi2007water}. Starting from the concept of Kalman filters, state space models \cite{hyndman2008forecasting} relate the measured values over time to the evolution of (unknown) states of the considered system. A comparison of the performance of ARIMA and state space models in a retail setting can be found, e.g., in \cite{ramos2015performance}. Structural time series models \cite{StructuralTS} take a similar approach and decompose the observations into components for trend, seasonality, and others. A modern implementation is Facebook's ``Prophet''  \cite{taylor2018forecasting}. Several other approaches are known in the literature as well: K\"ok and Fisher \cite{kok2007demand} model the demand based on the average demand for a product, modified by a logistic regression for the number of customers in a store, purchase incidence, and product choice. Wang \etal \cite{wang2005demand} model the demand under the assumption of a Poisson process.

% ML
Alternatively to the methods discussed above, time series forecasts in general and demand predictions in particular can be approached using supervised machine learning methods, using the values of the time series to be predicted as target, i.e., the numbers the machine learning algorithm is trained to predict.  For example, see \cite{Zhang2012,remus2001neural} for applications of neural networks for time series forecasting. Machine learning approaches can include multiple covariates. This allows to include not only past sales data, but also information about prices, promotions, or details about the article or store in the forecast. There are two main approaches for time series forecasting using  machine learning: We either assume that the variables are independent random variables (for both prediction target and potentially time-dependent features) for each time step of the different time series, or we consider the full sequences of time steps as a whole, resulting in a single sample (including prediction target and potentially time-dependent features) for each individual time series. For the case of individual samples for different time steps, in order to combine the endogenous information from the target autocorrelation with all other exogenous features, the most prominent approach is to include lagged target information explicitly via stacking of univariate time series predictions for the target as features in the machine learning model. For example, we could add exponentially smoothed moving averages with a range of dampening factors as features into the model. An example for the estimation of future demand in the context of price optimization, using regression trees \cite{breiman1984}, can be found in \cite{ferreira2016analytics}. For the case of sequence samples, deep learning methods such recurrent neural networks (RNN) \cite{rnn}, especially in form of ``long short-term memory'' (LSTM) networks \cite{hochreiter1997long}, or transformers \cite{transformer}, based on the self-attention mechanism, are used. These methods use information about the prediction target autocorrelation implicitly via the input sequences of the time-ordered training events. Both concepts, LSTM cells and self-attention, aim at learning from the data the elements of the individual time series that are deemed important for the prediction of the future values by the algorithm, and implicitly rely on the autocorrelation between the observed events at different times as well. The interest in deep learning approaches for demand forecasting has increased significantly recently, see, e.g., \cite{bandara2019sales,yu2017application, goyal2018solution,helmini2019sales, golkabek2020demand}. A review of deep learning methods for time series forecasting in a wide range of applications can also be found in \cite{langkvist2014review}, the use of RNNs with focus on industrial applications is discussed in \cite{dixon2020industrial}. A further approach is to explore the use of generative adversarial networks \cite{goodfellow2014generative} in the context of time series forecasting \cite{haas2020statistical, ramponi2018t, smith2020conditional}.

% temporal confounding
In many situations, relying on the exploitation of autocorrelation, either via traditional time series forecasting or machine learning approaches, is sufficient. However, this can lead to shortcomings in terms of explainability and generalizability due to temporal confounding. There are two main paths to avoid this: One way is to explicitly account for the confounders and extend the model to include them. The main challenge is then the discovery of temporal confounders and, for example, try to learn the causal structure from the observational data, see, e.g., \cite{malinsky2018causal,runge2018causal,Runge2019}. This topic is of particular importance in fields like medicine \cite{bica2020time} or climate research \cite{perrakis2014controlling}. Brodersen \etal \cite{Brodersen2015,GoogleCausalImpact} propose to use Bayesian structural time series models to estimate the causal effect of interventions on a time series by creating a counterfactual prediction that describes the behavior of the time series had the intervention not taken place. While structural time series equations allow to model components such as trend, seasonality, and others explicitly, the fundamental form of the state equation is $Y_t = G_t Y_{t-1} + \epsilon_t$, where $\epsilon_t \sim \mathcal{N}(0,\sigma^2)$ is a white noise term and $G_t$ some matrix that describes the evolution from the state at $Y_{t-1}$ to $Y_t$. Therefore, great care must be taken to avoid temporal confounding when setting up the states for structural time series models. Compared to fields like medicine, marketing, and many others that focus on the causal impact of few or just one intervention(s), retailers have to include many external effects that affect the predicted demand and are repeated frequently: Prices are changed dynamically, advertisements placed, new products are included in the assortment and others removed, etc.

@@

% quantile regression
Generally, quantile regression \cite{koenker2001,wen2017} can be implemented in various frameworks and used to estimate a range of quantiles for each predicted distribution, from which an empirical probability distribution can be interpolated. Using a dedicated neural network \cite{Feindt2006190}, either the full PDF or a defined range of quantiles can be calculated directly from the data for each individual prediction without assuming an underlying model. Focusing on time series, Hyndman \cite{hyndman1995highest} proposes to use either a simulation or bootstrapping framework to estimate the highest density forecast regions of non-normal time series. Tay \cite{tay2000density} summarizes the use of density forecasts in applications in finance and economics. Using RNNs, quantile regression for multiple forecasting horizons \cite{wen2017multi} has been used to predict the demand at the internet retailer Amazon. Using an attention-based architecture, the Temporal Fusion Transformer (TFT) \cite{lim2019temporal} can also predict a set of specified quantiles at each predicted time step. Another possibility to model the underlying data distribution without assumption of a fixed distribution class is via conditional normalizing flows \cite{rezende2016variational,timeseriesflow}, which aim to transform a simple initial density into a more complex one by applying a sequence of invertible transformations.

 % PDF assumption
Alternatively, we can start from a model assumption for the demand distribution and fit the model parameters instead of reconstructing the complete distribution \cite{astonpr373, SALINAS20201181}. This approach is computationally favorable and usually more robust, as fewer parameters need to be estimated. Empirically, we can determine the best fitting distribution from data \cite{adan1995}. However, given the stochastic nature of the demand, such an empirically determined distribution is not expected to be stable and prone to sudden changes. Instead, the choice of the demand distribution should be motivated by theoretic considerations. The discrete demand is typically modeled as a negative binomial distribution (NBD), also known as Gamma-Poisson distribution \cite{Ehrenberg1959, Ehrenberg1967, Ehrenberg1972, Chatfield1973, Schmittlein_1985}. This distribution arises if the Poisson parameter $\mu$ is a random variable itself, which follows a Gamma distribution. The NBD has two parameters, the mean $\mu$ and the variance $ \sigma^2 > \mu$, and is over-dispersed compared to the Poisson distribution, for which $\sigma^2 = \mu$. Hence, for each ordering decision, the model parameters $\mu$ and $\sigma^2$ need to be determined at the required granularity, typically for each item, sales location, and ordering time, depending on all auxiliary data describing article details, retail location, and influencing factors such as pricing and advertisement information.

@@

In the following, after a brief recap of the fundamental ideas of Cyclic Boosting, we describe a method to predict mean and variance for individual NBD using two Cyclic Boosting models.

@@

The main idea of this algorithm is the following: First, each feature, denoted by index $j$, is discretized appropriately into $k$ bins to reflect the specific behavior of the feature. The global mean $\mu$ is determined from all values $y$ of the  target variable $Y \in [0,\infty)$ observed in the data. Single data records, for example, the sales corresponding to a specific product-location-date combination along with all relevant features, are indexed by $i$.
The individual predictions $\hat{y_i}$  can then be calculated as:
\begin{equation} \label{eqn:cb}
\hat{y}_i = \mu \cdot \prod \limits_{j=1}^p f^k_j \quad \text{with}\; k=\{ x_{j,i} \in b^k_j\}
\end{equation}
The factors $f^k_j$ are the model parameters that are determined iteratively from the features until the algorithm converges. During training, regularization techniques are applied to avoid overfitting and improve the generalization ability of the algorithm. The deviation of each factor from $f^k_j=1$ can then be used to explain how a specific feature contributes to each individual prediction.

>>>>

Demand forecasting is a central component of the replenishment process for retailers, as it provides crucial input for subsequent decision making like ordering processes. In contrast to point estimates, such as the conditional mean of the underlying probability distribution, or confidence intervals, forecasting complete probability density functions allows to investigate the impact on operational metrics, which are important to define the business strategy, over the full range of the expected demand. Whereas metrics evaluating point estimates are widely used, methods for assessing the accuracy of predicted distributions are rare, and this work proposes new techniques for both qualitative and quantitative evaluation methods. Using the supervised machine learning method ``Cyclic Boosting'', complete individual probability density functions can be predicted such that each prediction is fully explainable. This is of particular importance for practitioners, as it allows to avoid ``black-box'' models and understand the contributing factors for each individual prediction. Another crucial aspect in terms of both explainability and generalizability of demand forecasting methods is the limitation of the influence of temporal confounding, which is prevalent in most state of the art approaches.

@@

% specific quantile
One of the main business operations of retailers is to ensure that sufficient goods are available in the store for customers to buy. This entails two parts: In a first step, we need to estimate the future customer demand to be able to judge how many goods we will need to order from the supplier or wholesaler, given the current inventory level. In the second step, aimed at inventory control, we need to place the actual orders, which is further complicated by the fact that multiple operational constraints, such as lot-sizes or delivery schedules, need to be included in the final ordering decision. Unfortunately, from the perspective of the retailer, the two parts, estimating the future demand and optimization of the order quantity, do not factorize into two separate issues, but are inextricably linked. This is because the business strategy depends on the complete chain of demand estimation and subsequent replenishment process to control the inventory. We can understand this in the following way: In the simplest case, a retailer might want to optimize their profit, or equivalently the cost, by reducing the impact of perishable items that have not been sold at the end of the selling period, as well as the lost revenue by not being able to meet the customer demand within this period. This is known as the ``newsvendor problem'' \cite{Edgeworth}, see, e.g., \cite{Khouja1999537} for a detailed review. Assuming linear overage ($h$) and underage ($b$) cost, the optimal quantile $q_{\mathrm{opt}} = {b}/{(b+h})$ of a known probability density function (PDF) describing demand as random variable can be calculated exactly. This quantile then corresponds to the numerical value of the demand PDF prediction that can be used to determine the optimal order quantity.

% simulations
In reality, the situation is often more complex, as many factors may contribute to the cost function. Furthermore, the cost function may vary between individual stock-keeping units (SKU) and may also change with time. However, the general principle remains: If we have access to the relevant cost function, we can determine the optimal point estimator of the future demand from the associated PDF, what is demonstrated in Appendix \ref{sec:CostQuantile}. For the case of a quadratic cost function, such as the mean squared error (MSE), or the linear cost function used in the newsvendor setting, corresponding to the mean absolute deviation (MAD), the calculations are straight-forward. If the cost function becomes more complex, we may no longer be able to perform the calculations analytically, but will have to resort to numerical approaches. Furthermore, if we can predict more than one quantile, we can investigate the effect of choosing different points, and determine, for example, how sensitive operational metrics, such as stockout or waste rate of perishable goods, change when choosing a different quantile. This can be done using, for example, a business oriented simulation that starts from the quantiles of the predicted demand for each SKU to simulate the value of the business relevant KPIs for each setting, taking operational constraints into account. This can then be used, for example, to align the business strategy. Predicting the full demand PDF therefore allows the maximum flexibility for further optimizations, because we have direct access to all quantiles.

% PDF predictions
Due to its stochastic nature, demand is difficult to forecast: It depends on many influencing factors and, as already stated above, we can interpret it as a random variable with associated PDF. And as such, demand is neither identically and independently distributed (i.i.d.) between different products and sales locations, nor between different times. This means that, while we can generally assume that demand follows a specific type of PDF, its parameters are unique to the instance for which an estimate is required. For example, the PDF describing the demand of a particular product-location-date combination is specific to the product, date, and retail location for which the forecast is made. Furthermore, this distribution depends on a wide range of influencing factors, such as seasonal effects, sales price, local weather, etc. Therefore, in order to predict the expected demand for each SKU at each sales location and sales day, we need to forecast individual PDFs rather than globally adjusted parameters for groups of SKUs or locations or sales days.

% temporal confounding
If we account for the cases where demand exceeds the current stock level, resulting in out-of-stock situations, the actual demand in the past can be approximated by realized sales. We can then understand the time-ordered observations of sales events as a time series of a random variable. It is therefore tempting to use established time series regression models to predict the future values of the expected demand. However, if we wish to use the predicted values for anything but the prediction itself, we have to be careful to avoid, or at least limit as much as possible, temporal confounding, which emerges if some external variable $X$ causally influences several time steps $t$ of the observed time series variable $Y$, say $Y_{t}$ and $Y_{t-\mathrm{lag}}$, where lag denotes a specific time difference between the two steps. In the case of a retailer, $Y_{t}$ is the demand for a specific product at a given sales location for a specific day, and $X$ can be, for example, the price of the product which we can set as part of our business strategy or change according to some promotion. This dependency creates a backdoor path between $Y_{t-\mathrm{lag}}$ and $Y_{t}$, that in turn leads to a spurious autocorrelation between the values of $Y$ at different times $t$, if we cannot control for the confounder $X$. This is particularly important for the case of demand forecasting, since customer demand is not a mere natural phenomenon we observe as an outsider, but retailers actively influence the customer demand  through a wide range of interventions, such as a given assortment choice, promotions, advertisements, coupons, etc. If we cannot disentangle the fundamental causal dependencies from the spurious autocorrelation, understanding the influences that lead to a specific prediction becomes very challenging, even if the underlying prediction method itself is fully explainable. Moreover, due to this disguising of actual causal dependencies, reliance on temporal confounding is also detrimental in terms of generalizability of the forecasting model. The issue of temporal confounding, together with a proposed solution, is discussed in more detail in sec. \ref{sec:ts_confounding}.

% ML
Demand forecasting for retailers often requires to predict millions of product-location combinations each day. Univariate time series forecasting methods operate on each of these many individual time series separately. On the other hand, supervised machine learning (ML) algorithms optimize all individual time series together, which can enhance the generalizability of the method by exploiting common aspects between the various time series. This can lead to a drastic reduction in variance of the model, and in turn to an improvement of the forecast quality of each of the individual time series. Furthermore, machine learning provides a straight-forward way to include the many exogenous variables influencing demand as features (or covariates) of the model. This can also help to improve the forecast quality compared to methods mainly relying on autocorrelation of the historic sales.

Hence, in order for a retailer to be able to use demand forecasts most effectively, we need to be able to:
\begin{itemize}
	\item {Predict the future demands on SKU level at required temporal granularity as complete individual PDFs.}
	\item {Avoid temporal confounding in the model used for the prediction of the future demand.}
	\item {Evaluate that each part of the PDF models the observed data appropriately.}
	\item {Ideally, each prediction is fully explainable, enabling operational experts to exactly understand what contributed to it.}
\end{itemize}

The remainder of the paper is organized as follows: We first review the relevant literature and existing work in sec. \ref{sec:LitRev}. We then describe our method to predict individual negative binomial PDFs using  a parametric approach including two distinct machine learning models for mean and variance in sec. \ref{sec:pdfEstimation}. After that, we describe novel techniques for the qualitative and quantitative evaluation of PDF predictions in sec. \ref{sec:pdfEvaluation}. This is followed by a detailed discussion on temporal confounding in sec. \ref{sec:ts_confounding}. And finally, we present a demand forecasting example to show an application of our methods in sec. \ref{sec:example}.

@@

% univariate time series methods
A popular approach is to treat the demand as a time series, i.e., as a time-ordered sequence of observed sales events. In this setting, the future demand can be predicted as a point estimate using auto-regressive integrated moving average (ARIMA) \cite{BoxJenkins} or exponential smoothing \cite{croston1972forecasting, holt1957forecasting, Brown1963, Gardner1985} approaches. See also \cite{deGooijerHyndman2006} for a comprehensive overview over traditional time series forecasting methods. Essentially, these methods extrapolate the future values by building some form of regression model based on past observations, i.e., $y_{t+1} = f(y_t, y_{t-1}, \ldots)$, introducing dedicated functional components for trends, seasonality, or external factors in various ways, depending on the exact modeling approach taken. The main underlying concept behind these approaches is to exploit the autocorrelation between values of the variable $Y$ at different times $t$. Intuitively, we assume that  the sequence of events observed so far is a good predictor of future events. Examples for these methods in the context of demand forecasting can be found in, e.g., \cite{huber2017cluster, kalchschmidt2006forecasting, fattah2018forecasting, permatasari2018sales}.  A different approach is to use a concept from engineering, the Kalman filter \cite{kalman1960new}, and apply it to statistical forecasting \cite{morrison1977kalman}. Examples for this method used to estimate demand are, e.g., \cite{mitropoulos1980using, tegene1991kalman, kandananond2014applying,jacobi2007water}. Starting from the concept of Kalman filters, state space models \cite{hyndman2008forecasting} relate the measured values over time to the evolution of (unknown) states of the considered system. A comparison of the performance of ARIMA and state space models in a retail setting can be found, e.g., in \cite{ramos2015performance}. Structural time series models \cite{StructuralTS} take a similar approach and decompose the observations into components for trend, seasonality, and others. A modern implementation is Facebook's ``Prophet''  \cite{taylor2018forecasting}. Several other approaches are known in the literature as well: K\"ok and Fisher \cite{kok2007demand} model the demand based on the average demand for a product, modified by a logistic regression for the number of customers in a store, purchase incidence, and product choice. Wang \etal \cite{wang2005demand} model the demand under the assumption of a Poisson process.

% ML
Alternatively to the methods discussed above, time series forecasts in general and demand predictions in particular can be approached using supervised machine learning methods, using the values of the time series to be predicted as target, i.e., the numbers the machine learning algorithm is trained to predict.  For example, see \cite{Zhang2012,remus2001neural} for applications of neural networks for time series forecasting. Machine learning approaches can include multiple covariates. This allows to include not only past sales data, but also information about prices, promotions, or details about the article or store in the forecast. There are two main approaches for time series forecasting using  machine learning: We either assume that the variables are independent random variables (for both prediction target and potentially time-dependent features) for each time step of the different time series, or we consider the full sequences of time steps as a whole, resulting in a single sample (including prediction target and potentially time-dependent features) for each individual time series. For the case of individual samples for different time steps, in order to combine the endogenous information from the target autocorrelation with all other exogenous features, the most prominent approach is to include lagged target information explicitly via stacking of univariate time series predictions for the target as features in the machine learning model. For example, we could add exponentially smoothed moving averages with a range of dampening factors as features into the model. An example for the estimation of future demand in the context of price optimization, using regression trees \cite{breiman1984}, can be found in \cite{ferreira2016analytics}. For the case of sequence samples, deep learning methods such as recurrent neural networks (RNN) \cite{rnn}, especially in form of ``long short-term memory'' (LSTM) networks \cite{hochreiter1997long}, or transformers \cite{transformer}, based on the self-attention mechanism, are used. These methods use information about the prediction target autocorrelation implicitly via the input sequences of the time-ordered training events. Both concepts, LSTM cells and self-attention, aim at learning from the data the elements of the individual time series that are deemed important for the prediction of the future values by the algorithm, and implicitly rely on the autocorrelation between the observed events at different times as well. The interest in deep learning approaches for demand forecasting has increased significantly recently, see, e.g., \cite{bandara2019sales,yu2017application, goyal2018solution,helmini2019sales, golkabek2020demand}. A review of deep learning methods for time series forecasting in a wide range of applications can also be found in \cite{langkvist2014review}, the use of RNNs with focus on industrial applications is discussed in \cite{dixon2020industrial}. A further approach is to explore the use of generative adversarial networks \cite{goodfellow2014generative} in the context of time series forecasting \cite{haas2020statistical, ramponi2018t, smith2020conditional}.

% temporal confounding
In many situations, relying on the exploitation of autocorrelation, either via traditional time series forecasting or machine learning approaches, is sufficient. However, this can lead to shortcomings in terms of explainability and generalizability due to temporal confounding. There are two main paths to avoid this: One way is to explicitly account for the confounders and extend the model to include them. The main challenge is then the discovery of temporal confounders and, for example, try to learn the causal structure from the observational data, see, e.g., \cite{malinsky2018causal,runge2018causal,Runge2019}. This topic is of particular importance in fields like medicine \cite{bica2020time} or climate research \cite{perrakis2014controlling}. Brodersen \etal \cite{Brodersen2015,GoogleCausalImpact} propose to use Bayesian structural time series models to estimate the causal effect of interventions on a time series by creating a counterfactual prediction that describes the behavior of the time series had the intervention not taken place. While structural time series equations allow to model components such as trend, seasonality, and others explicitly, the fundamental form of the state equation is $Y_t = G_t Y_{t-1} + \epsilon_t$, where $\epsilon_t \sim \mathcal{N}(0,\sigma^2)$ is a white noise term and $G_t$ some matrix that describes the evolution from the state $Y_{t-1}$ to $Y_t$. Therefore, great care must be taken to avoid temporal confounding when setting up the states for structural time series models. Compared to fields like medicine, marketing, and many others that focus on the causal impact of few or just one intervention(s), retailers have to include many external effects that affect the predicted demand and are repeated frequently: Prices are changed dynamically, advertisements placed, new products are included in the assortment and others removed, etc.

@@

% quantile regression
Generally, quantile regression \cite{koenker2001} can be implemented in various frameworks and used to estimate a range of quantiles for each predicted distribution, from which an empirical probability distribution can be interpolated. Using a dedicated neural network \cite{Feindt2006190}, either the full PDF or a defined range of quantiles can be calculated directly from the data for each individual prediction without assuming an underlying model. Focusing on time series, Hyndman \cite{hyndman1995highest} proposes to use either a simulation or bootstrapping framework to estimate the highest density forecast regions of non-normal time series. Tay \cite{tay2000density} summarizes the use of density forecasts in applications in finance and economics. Using RNNs, quantile regression for multiple forecasting horizons \cite{wen2017multi} has been used to predict the demand at the internet retailer Amazon. Using an attention-based architecture, the Temporal Fusion Transformer (TFT) \cite{lim2019temporal} can also predict a set of specified quantiles at each predicted time step. Another possibility to model the underlying data distribution without assumption of a fixed distribution class is the usage of conditional normalizing flows \cite{rezende2016variational,timeseriesflow}, which aim to transform a simple initial density into a more complex one by applying a sequence of invertible transformations.

 % PDF assumption
Alternatively, we can start from a model assumption for the demand distribution and fit the model parameters instead of reconstructing the complete distribution \cite{astonpr373, SALINAS20201181}. This approach is computationally favorable and usually more robust, as fewer parameters need to be estimated. Empirically, we can determine the best fitting distribution from data \cite{adan1995}. However, given the stochastic nature of demand, such an empirically determined distribution is not expected to be stable and prone to sudden changes. Instead, the choice of the demand distribution should be motivated by theoretic considerations. Discrete demand is typically modeled as a negative binomial distribution (NBD), also known as Gamma-Poisson distribution \cite{Ehrenberg1959, Ehrenberg1967, Ehrenberg1972, Chatfield1973, Schmittlein_1985}. This distribution arises if the Poisson parameter $\mu$ is a random variable itself, which follows a Gamma distribution. The NBD has two parameters, the mean $\mu$ and the variance $ \sigma^2 > \mu$, and is over-dispersed compared to the Poisson distribution, for which $\sigma^2 = \mu$. Hence, for each ordering decision, the model parameters $\mu$ and $\sigma^2$ need to be determined at the required granularity, typically for each item, sales location, and ordering time, depending on all auxiliary data describing article details, retail location, and influencing factors such as pricing and advertisement information.

@@

In the following, after a brief recap of the fundamental ideas of Cyclic Boosting, we describe a method to predict mean and variance for individual NBDs using two Cyclic Boosting models.

@@

The main idea of this algorithm is the following: First, each feature, denoted by index $j$, is discretized appropriately into $k$ bins to reflect the specific behavior of the feature. The global mean $\mu$ is determined from all values $y$ of the  target variable $Y \in [0,\infty)$ observed in the data. Single data records, for example, the sales corresponding to a specific product-location-date combination (meaning the sales record of a specific item sold on a specific day at a specific sales location) along with all relevant features, are indexed by $i$.
The individual predictions $\hat{y_i}$  can then be calculated as:
\begin{equation} \label{eqn:cb}
\hat{y}_i = \mu \cdot \prod \limits_{j=1}^p f^k_j \quad \text{with}\; k=\{ x_{j,i} \in b^k_j\}
\end{equation}
The factors $f^k_j$ are the model parameters that are determined iteratively from the features until the algorithm converges. During training, regularization techniques are applied to avoid overfitting and improve the generalization ability of the algorithm. The deviation of each factor from $f^k_j=1$ can then be used to explain how a specific feature contributes to each individual prediction.

In detail, the following meta-algorithm describes how the model parameters $f^k_j$ are obtained from the training data:
\begin{enumerate}
\item{Calculate the global average $\mu$ from all observed $y$ across all bins $k$ and features $j$.}
\item{Initialize the factors $f^k_j \leftarrow 1$}
\item{Cyclically iterate through features $j = 1,...,p $ and calculate in turn for each bin $k$ the partial factors $g$ and corresponding aggregated factors $f$, where indices $t$ (current iteration) and $\tau$ (current or preceding iteration) refer to iterations of full feature cycles as the training of the algorithm progresses:
\begin{equation} \label{factors}
g^k_{j,t} = \frac{\sum \limits_{x_{j,i} \in b^k_j} y_i}{\sum \limits_{x_{j,i} \in b^k_j} \hat{y}_{i,\tau}}\;\; \mathrm{where} \; \; f^k_{j,t} = \prod \limits_{s=1}^t g^k_{j,s}
\end{equation}
Here,  $g$ is a factor that is multiplied to the corresponding $f_{t-1}$ in each iteration. The current prediction, $\hat{y}_\tau$, is calculated according to \eqn \eqref{eqn:cb} with the current values of the aggregated factors $f$:
\begin{equation} \label{factors3}
\hat{y}_{i,\tau} = \mu \cdot \prod \limits_{j=1}^p f^k_{j,\tau}
\end{equation}
To be precise, the determination of $g^k_{j,t}$ for a specific feature $j$ uses $f^k_{j,t-1}$ in the calculation of $\hat{y}$. For the factors of all other features, the newest available values are used, i.e., depending on the sequence of features in the algorithm, either from the current ($\tau=t$) or the preceding iteration ($\tau=t-1$).}
\item{Quit when stopping criteria, e.g., the maximum number of iterations or no further improvement of an error metric such as the MAD or MSE, are met at the end of a full feature cycle.}
\end{enumerate}
